Effectiveness of BNT162b2 booster doses in England: a observational study in OpenSAFELY

### 2022-02-22 manuscript/draft-manuscript.Rmd
addressing comments

flowchart <- read_csv(fs::path(output_dir_os, "flowchart.csv"))
cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
matchsummary_treated <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 1L)
matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
incidences <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0)

@@

We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls with only two doses. During days 14-20 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death.

@@

<!-- **Keywords** COVID-19; Vaccine effectiveness; Target trial; EHR data -->

@@

<!-- * flowchart incl matching -->

@@

<!-- * JCVI group spell out -->

@@

A total of `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment group. Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1).

@@

Participant characteristics at trial recruitment between treatment groups (Table 1) were typically well-balanced (necessarily so on matched characteristics), though with some exceptions, for instance unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency. These unbalanced characteristics were adjusted for in the Cox regression models.

@@

```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
incidences %>%
  filter(
    fup_period == "All",
    outcome %in% c("postest", "covidadmitted", "coviddeath"),
    treatment == "pfizer"
  ) %>%
  left_join(
    matchsummary %>% select(treatment, n_eligible, prop_matched_eligible),
    by=c("treatment")
  ) %>%
  mutate(
    treatment_decr = paste0(
      treatment_descr, " - ",
      label_number(1, big.mark=",")(n_eligible), "eligible, ",
      label_number(1, big.mark=",")(n_1), "matched (",
      label_percent(0.1)(prop_matched_eligible), ")."
    )
  ) %>%
  select(
    treatment_descr, 
    outcome_descr,
    q_1,
    rate_1,
    q_0, 
    rate_0,
    rrECI
  ) %>%
  mutate(
    rate_1=if_else(rate_1<0.001, "<0.001", label_number(0.001)(rate_1)),
    rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
  )
  gt(
    groupname_col = c("treatment_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    outcome_descr = "Outcome",

@@

In the 14-20 days period of follow-up, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).

<!-- `r filter(hazards_none, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 A&E attendances, `r filter(hazards_none, outcome=="covidcc", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 critical care admission,  -->

<!-- For mRNA-1273, the corresponding estimated effectiveness were `r filter(hazards_none, outcome=="postest", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 A&E attendances was `r filter(hazards_none, outcome=="covidattendance", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 hospital admission was `r filter(hazards_none, outcome=="covidadmitted", treatment=="moderna", term_left==14) %>% pull(ve_display)`, for COVID-19 critical care admission was `r filter(hazards_none, outcome=="covidcc", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, and for COVID-19 death was `r filter(hazards_none, outcome=="coviddeath", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`.  -->

For the same period, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only separately for 2× BNT162b2 and 2× ChAdOx1-S primary course recipients respectively were `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).

>>>>

flowchart <- read_csv(fs::path(output_dir_os, "flowchart.csv"))
cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
matchsummary_treated <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 1L)
matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
matchflowchart <- read_csv(fs::path(output_dir_os, "matchflowchart.csv"))
matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
incidences <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0)

@@

We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls with only two doses. During days 14-20 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death.

@@

**Keywords** COVID-19; Booster vaccine effectiveness; Target trial; EHR data

@@

@@

@@

A total of `r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` BNT162b2 booster recipients were identified between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion into the treatment group. Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure S1).

@@

Participant characteristics at trial recruitment between treatment groups (Table 1) were typically well-balanced though with some exceptions. For instance, unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency.

@@

```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
incidences %>%
  filter(
    fup_period == "All",
    outcome %in% c("postest", "covidadmitted", "coviddeath"),
    treatment == "pfizer"
  ) %>%
  left_join(
    matchsummary %>% select(treatment, n_eligible, prop_matched_eligible),
    by=c("treatment")
  ) %>%
  mutate(
    treatment_decr = paste0(
      treatment_descr, " - ",
      label_number(1, big.mark=",")(n_eligible), "eligible, ",
      label_number(1, big.mark=",")(n_1), "matched (",
      label_percent(0.1)(prop_matched_eligible), ")."
    )
  ) %>%
  select(
    treatment_descr, 
    outcome_descr,
    q_1,
    rate_1,
    q_0, 
    rate_0,
    rrECI
  ) %>%
  mutate(
    rate_1=if_else(rate_1<0.001, "<0.001", label_number(0.001)(rate_1)),
    rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
  ) %>%
  gt(
    groupname_col = c("treatment_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    outcome_descr = "Outcome",

@@

In the 14-20 days period of follow-up, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).

<!-- `r filter(hazards, subgroup=="none", outcome=="covidattendance", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 A&E attendances, `r filter(hazards, subgroup=="none", outcome=="covidcc", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 critical care admission,  -->

<!-- For mRNA-1273, the corresponding estimated effectiveness were `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 A&E attendances was `r filter(hazards, subgroup=="none", outcome=="covidattendance", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 hospital admission was `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="moderna", term_left==14) %>% pull(ve_display)`, for COVID-19 critical care admission was `r filter(hazards, subgroup=="none", outcome=="covidcc", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, and for COVID-19 death was `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`.  -->

For the same period, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only separately for 2× BNT162b2 and 2× ChAdOx1-S primary course recipients respectively were `r filter(hazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
